---
title: "funWithTwitter Overview"
author: "Omar AlOmeir"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

## Package Overview

The main purpose of this package is to provide functions that connect to the twitter api and allow the user to explore and embed tweets into their R markdown reports. There are functions to display tweets using tweets' IDs either copied from a tweet's URL, for example: "https://twitter.com/STAT545/status/506900131086880768" or acquired using the function `getUserTimeline()`. `getUserTimeline()` is used to acquire a number of the most recent tweets in a user's timeline. The IDs can be used to display the tweet as text using `getTweet()` or wrapped in html to embed in Rmarkdown reports using `embedTweet()`.

## Functions

First load the package:
```{r}
library(funWithTwitter)
```

### getUserTimeline()

This is the main function from the package. This is a function to return tweet ids from a twitter user's timeline starting from the most recent tweet.
The results can be used as input for `getTweet()` or `embedTweet()` functions.

Examples:
```{r}
(getUserTimeline(screen_name = "STAT545"))

(getUserTimeline(screen_name = "STAT545", include_rts = FALSE))

(getUserTimeline(screen_name = "STAT545", count = 5, exclude_replies = FALSE, include_rts = FALSE))
```

### getTweet()

This function is used to display a tweet's text along with the screen_name of the user who wrote it. The function takes a tweet's id which can be found in a tweet's URL. Example: "https://twitter.com/STAT545/status/506900131086880768"

```{r}
getTweet()

getTweet("535530698451726336")
```

### embedTweet()

This function returns a tweet's text, the username, and the user's avatar wrapped in html code to be used in R markdown reports. The chunk needs to have the options `results='asis'` for this to work.

```{r results='asis'}
embedTweet()

embedTweet("535530698451726336")
```

### Futher use

The `getUserTimeline()` function can be used with `getTweet()` or `embedTweet()` for some interesting usecases. For example:

* Print the most recent tweet from the "STAT545" account:
```{r}
tweet <- getUserTimeline("STAT545", count = 1)
(getTweet(tweet))
```

* Print the most recent tweet that is not a retweet:
```{r}
tweet <- getUserTimeline("STAT545", count = 1, include_rt = FALSE)
(getTweet(tweet))
```

For multiple tweets we can use the a*ply family function from `plyr` to do some interesting things. For example:

* Print the 5 most recent tweets:
```{r results='asis'}
tweets <- getUserTimeline("STAT545", count = 5)

knitr::kable(plyr::adply(tweets, 1, getTweet))
```

* Embed the 5 most recent tweets:
```{r results='asis'}
tweets <- getUserTimeline("STAT545", count = 5)

plyr::a_ply(tweets, 1, embedTweet)
```

## Workflow and Reflections

* The user could face some issues with the twitter api as it does not work as expected when asking for the timeline of a certain user. The `count` variable is not exactly the number of tweets you would get if you exclude retweets, it would only return the non-retweets within the last n tweets.
* There was a number of things that proved difficult to do, the authentication used by twitter has changed a lot since the last time I used it. This is not exactly encouraging for app developers.
* I started out by making this package as a wrapper of sorts for the twitteR package. Unfortunatley, the twitteR version in CRAN did not work, and I had no idea how to import the development version of a package into my package. Instead I went for the `httr` package to handle http GET requests and twitter authentication and it works without obscuring any important details.
* The authentication works wihtout hiding the consumer or token secrets. This is not exactly ideal but there is no other way to do it in an open source package that does not invlove too much work. The app I'm using here is not really important and it's linked to a newly created twitter account, so I though instead of trying to solve a well known oauth problem I'll just leave it as is.
* There's a lot I could add to improve this package; the css can be made external instead of internal in the `embedTweet` function, the `getUserTimeline` function could use more options, finally I could try to add more functions to search and collect tweets. However, for our purposes what is there is suffecient and users can always manually collect tweets ids and then embed them.
* Last minute error, for some reason the vignette would fail the `devtools::check()` without the `library(funWithTwitter)` call in the vignette.
